---
title: "Tutorial-04"
# author: Kate Saunders
format: 
  html:
    toc: true
    css: ../../assets/tutorial.css
    embed-resources: true
  pdf:
    toc: true
editor: visual
---

## Visualisation in R

### Overview 

Today you will be learning how to create visualisations in R. This will involve adding layers to a plot using the grammar of graphics. 

### Learning Objectives

-  Develop the skills to create plots in R

-  Develop your understanding of the grammar of graphics 

- Learn about different plot geometries 

- Learn about different ways to encode colour

- Practice changing the plot appearance (e.g. theme)

### Preparation

-   Ensure you installed R and RStudio 

-   Ensure you have installed the `ggplot2` package. You can install this package directly, but this package is also part of the `tidyverse` package.

-   Set yourself an R project for this tutorial.

**If you are confused about any of the above, let us know! It's important you don't get left behind here and we want to help you.** You may also like to refer back to the Lecture 1 and Tutorial 1 material. 

## Tasks

### Exercise 1 

Today you will be creating visualisations in `ggplot2` to analyse sporting statistics from the Boston Celtics NBA basketball team.

```{r echo = FALSE}
show_solutions = FALSE
```

1.  Download the dataset from Moodle, `boston_celtics.csv,` and place it in a folder called `data` within the R Project you set up. Then read your data set into R.
  
We recommend installing the package `here` to help keep things organised when referencing files.

```{r echo = TRUE, warning = FALSE, message = FALSE}
if(!require(here)){
  install.packages("here")
}
library(here)
```

```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
boston_celtics <- read_csv(here("data", "boston_celtics.csv"))
```

We also recommend following the Reading Data Checklist if you have any problems (see lecture).

```{r echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE}

# Check your working directory
getwd()  

# Check your data file is where you think it is
file.exists(here("data","boston_celtics.csv"))

# Read in your data 
library(tidyverse)
boston_celtics <- read_csv(here("data", "boston_celtics.csv"))

# Look at your data 
View(boston_celtics)

# Look at a summary
summary(boston_celtics)
```

2.  Create a scatter plot that shows the `game_date` and `team_scores`. Make the colour of the points `forestgreen`. Focus on the geometry and aesthetic layer for now.

```{r eval = FALSE, echo = TRUE}
    ggplot(data = ???) + 
      geom_???(aes(x = ???, y = ???), col = ???)
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score), col = "forestgreen")
```

3.  Colour the points according to if the team won or lost. Use the hex colour codes from the [Boston Celtics team](%22https://teamcolorcodes.com/boston-celtics-color-codes/%22) to set the colour scheme. *Notice that the colour command now moves inside the bracket specify aesthetic mappings - Make sure you understand why this is.*

```{r eval = FALSE, echo = TRUE}
    ggplot(data = ???) + 
      geom_???(aes(x = ???, y = ???, col = ???))
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score, col = team_winner)) + 
  scale_colour_manual(label = c("Loss", "Win"), values = c("TRUE" = "#007A33", "FALSE" = "#BA9653"))
```

4.  As there are a lot of points, try changing the size and the transparency to see if that improves the plot.

```{r eval = FALSE, echo = TRUE}
ggplot(data = ???) + 
  geom_???(
    aes(x = ???, y = ???, col = ???), 
    size = ???, alpha = ???
    )
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score, col = team_winner), size = 0.8, alpha = 0.6) + 
  scale_colour_manual(label = c("Loss", "Win"), values = c("TRUE" = "#007A33", "FALSE" = "#BA9653"))
```

5.    Look at the [theme options](%22https://ggplot2.tidyverse.org/reference/ggtheme.html%22) and try using a few different ones, before deciding on the theme for your plot.

```{r eval = FALSE, echo = TRUE}
ggplot(data = ???) + 
  geom_???(
    aes(x = ???, y = ???, col = ???), 
    size = ???, alpha = ???
    ) + 
  theme_???()
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score, col = team_winner), size = 0.8, alpha = 0.6) + 
  scale_colour_manual(label = c("Loss", "Win"), values = c("TRUE" = "#007A33", "FALSE" = "#BA9653")) + 
  theme_bw()
```

5.  Add appropriate labels to your plot.

```{r eval = FALSE, echo = TRUE}
ggplot(data = ???) + 
  geom_???(
    aes(x = ???, y = ???, col = ???), 
    size = ???, alpha = ???
    ) + 
  theme_???() + 
  labs(title = ???, subtitle = ???, x = ???, y = ???)
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score, col = team_winner), size = 0.8, alpha = 0.6) + 
  scale_colour_manual(label = c("Loss", "Win"), values = c("TRUE" = "#007A33", "FALSE" = "#BA9653")) + 
  theme_bw() + 
  labs(title = "Boston Celtics Team Score", 
       subtitle = "2021 - Current Season", 
       x = "Date", 
       y = "Points")
```

6.  Move the legend on the plot to increase the data-density. *(Note the solutions show a more advanced way to move the legend than the lecture notes.)*

```{r eval = FALSE, echo = TRUE}
ggplot(data = ???) + 
  geom_???(
    aes(x = ???, y = ???, col = ???), 
    size = ???, alpha = ???
    ) + 
  theme_???() + 
  labs(title = ???, subtitle = ???, x = ???, y = ???)
  theme(legend.position = ???)
```

```{r echo = show_solutions, eval = show_solutions}
ggplot(data = boston_celtics) + 
  geom_point(aes(x = game_date, y = team_score, col = team_winner), size = 0.8, alpha = 0.6) + 
  scale_colour_manual(label = c("Loss", "Win"), values = c("TRUE" = "#007A33", "FALSE" = "#BA9653")) + 
  theme_bw() + 
  labs(title = "Boston Celtics Team Score", 
       subtitle = "2021 - Current Season", 
       x = "Date", 
       y = "Points") + 
  theme(legend.position = c(0.8, 1.1), legend.title = element_blank(), legend.direction = "horizontal")
```

### Exercise 2

Now we've mastered the basic layers, take some time to creating some other visualisations that are interesting to you!

-   You may like to produce plots looking at the distribution of some of the key variables. For example, `assists`, `rebounds`, `steals` etc.

```{r echo = show_solutions, eval = show_solutions}
boston_celtics |>
  ggplot(aes(x = assists)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

boston_celtics |>
  ggplot(aes(x = steals)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

boston_celtics |>
  ggplot() + 
  geom_histogram(aes(x = total_rebounds), fill = "blue", alpha = 0.3) +
  geom_histogram(aes(x = offensive_rebounds), fill = "green", alpha = 0.3) + 
  geom_histogram(aes(x = defensive_rebounds), fill = "orange", alpha = 0.3)
```

-   Maybe you are interested in how the shot percentage varies between `field_goal_pct` (2 pts) and `three_point_field_goal_pct` (3 pts).

```{r echo = show_solutions, eval = show_solutions}
boston_celtics |>
  # filter(season == 2025) |> 
  ggplot() + 
  geom_point(aes(x = field_goal_pct, y = three_point_field_goal_pct)) +
  geom_abline(slope = 1, linetype = "dotted") + 
  coord_fixed()
```

-   Perhaps you want to create a bar chart (`geom_bar`) to see how many times they've played opposing teams this season.

```{r echo = show_solutions, eval = show_solutions}
boston_celtics |>
  filter(season == 2025) |>
  ggplot() + 
  geom_bar(aes(y = opponent_team_name))
```

## Wrap Up

By the end of this tutorial you should feel comfortable reading in a data set into R and creating a visualisation by adding layers in `ggplot2`. 

**Important:** We will continue to build on this knowledge in coming weeks, so if you had any difficulties come along to consult to get some 1-on-1 help. We can slowly walk you through any parts you are struggling with so it makes sense.

## In Your Own Time 

Go through the code in exercise 2 line by line to check you understanding. 
You may like to also add labels and themes to these plots. Have a think about what else you might do to improve them. 

<!-- ##### Material developed by Dr. Kate Saunders. -->

##### © Copyright 2025 Monash University
