---
pagetitle: "ETX2250/ETF5922"
subtitle: "Visualisations in R: Part 2"
author: Kate Saunders
email: "etx2250-etf5922.caulfield-x@monash.edu"
date: "Lecture 6"
department: "Department of Econometrics and Business Statistics"
unit-url: "dvac.ss.numbat.space"
footer: "ETX2250/ETF5922"
format: 
  revealjs: 
    logo: images/monash-one-line-black-rgb.png
    slide-number: c
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: false
    controls: true
    width: 1280
    height: 720
    css: [../assets/custom.css, ../assets/lecture-01.css]
    include-after-body: "../assets/after-body.html"
    #   boardmarker-width: 5
    #   buttons: true
    embed-resources: true
---

```{r echo = FALSE, message=FALSE, warning = FALSE}

# if(!require(GGally))
#   install.packages("GGally")

if(!require(tidyverse))
  install.packages("tidyverse")

# if(!require(patchwork))
#   install.packages("patchwork")

# library(GGally)
library(tidyverse)
library(patchwork)
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etx2250-etf5922-title background-image="images/bg-01.png"}
### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">

<li><i class="fas fa-envelope"></i>`r rmarkdown::metadata$email`</li>

<li><i class="fas fa-calendar-alt"></i> `r rmarkdown::metadata$date`</li>

<li><i class="fa-solid fa-globe"></i>\<a href="`r rmarkdown::metadata[["unit-url"]]`"\>`r rmarkdown::metadata[["unit-url"]]`</a></li>

</ul>

<br>
:::

## What we've covered 

:::callout-important 
## We've learnt ... 

:::incremental

* We can map variables to different geometries (bar charts, line plots, scatter plots etc), and 

* We can map variables to different pre-attentive attributes (colour, shape, size, etc)

* We also know small multiples can be highly effective to show the data when there is too much information on a single plot

* Breaking a plot into panels by category can help!

::: 

::: 

## Today's Lecture

:::: callout-note
## Learning Objectives

::: incremental
-   Grouping 
    - `group` aesthetic mapping 
    
-   Small multiples in R aka. facetting 
    -   `facet_wrap()`
    -   `facet_grid()`
    
-   Combining plots together
    -   `patchwork` Package
    
:::
::::

<!-- -   Summary plots in higher dimensions -->
<!--     -   `ggpairs()` -->

# Groups {background-color="#006DAE"}

## Example Data 

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE

library(tidyverse)
energy = read_csv("data/energydata.csv")
weather = read_csv("data/weather.csv")
energy_weather = full_join(energy, weather, by = c("Date", "State"))
head(energy_weather)
```

## Let's plot it 

What is the relationship between Temperature and Demand? 

```{r}
#| message: FALSE
#| warning: FALSE
#| echo: TRUE
#| eval: FALSE 

ggplot(data = energy_weather,
       aes(x = MaxTemp, y = Demand)) + 
  geom_point() + 
  geom_smooth()
```

## 

:::callout-caution 
## Something weird is going on? 

* `geom_smooth()` is meant to show smoothed trend line 
* But this isn't what we would expect of this relationship 

::: 

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-align: center

ggplot(data = energy_weather, aes(x = MaxTemp, y = Demand)) + 
  geom_point() + 
  geom_smooth()
```

## Latent Variable

:::callout-tip 
## Found it 
`State` variable wasn't visualised! - Show using `col`

But - we need a trend for each `State`
:::

```{r}
#| message: FALSE
#| warning: FALSE
#| echo: FALSE
#| fig-align: center

ggplot(data = energy_weather,
       aes(x = MaxTemp, y = Demand)) + 
  geom_point(aes(col = State)) + 
  geom_smooth()
```

## Code for previous plot

```{r}
#| message: FALSE
#| warning: FALSE
#| echo: TRUE
#| eval: FALSE

ggplot(data = energy_weather,
       aes(x = MaxTemp, y = Demand)) + 
  geom_point(aes(col = State)) + 
  geom_smooth()
```

## Enter groups 

```{r}
#| message: FALSE
#| warning: FALSE
#| eval: TRUE
#| echo: true

ggplot(data = energy_weather, aes(x = MaxTemp, y = Demand)) + 
  geom_point(aes(col = State)) + 
  geom_smooth(aes(group = State))
```

## Even Better 

```{r}
#| echo: FALSE
#| fig-align: center

ggplot(data = energy_weather, 
       aes(x = MaxTemp, y = Demand, col = State, group = State)) + 
  geom_point() + 
  geom_smooth() 
```

## Your turn 

:::callout-caution 
## Your turn 

* Join the electricity and weather data together 

* Recreate the plot on the previous slide with coloured trend lines

* Think about which aesthetic mappings go in the top `ggplot()` call and which in the `geom` layers

:::

## Take a moment 

:::callout-note
## Reflect

* Does this plot show our data well? Could we do better?

:::

. . . 

::: callout-important 
## Some notes on `geom_smooth()`

* `geom_smooth()` can imply a trend or relationship that isn’t really there 

* Only use it if smoothing is appropriate for your data! 

:::

# Facetting {background-color="#006DAE"}

## Facetting

::: callout-tip
## Back to Tufte's Principles 

-   Sometimes we cannot display everything on a single plot

-   Small multiples are a way to display information on a plot efficiently

::: 

. . . 

:::callout-note 
## In R 

-   `facet_wrap`: wraps a sequence of panels based on one variable

-   `facet_grid`: forms a matrix of panels defined by row and column variables.

Let's see how this works 

:::

## Groups or facets

<br>

:::callout-tip
## Comparisons

* `facet_wrap` is useful to look at trends/patterns in each category 

* The `group` aesthetic is useful for comparing trends/patterns across categories in a single plot

::: 

. . . 

Example use: 

```{r}
#| warning: FALSE
#| message: FALSE
#| eval: FALSE
#| echo: TRUE
#| 
energy_weather |> 
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.4) + 
  geom_smooth() + 
  facet_wrap(vars(State))
```

## `facet_wrap`

```{r}
#| warning: FALSE
#| message: FALSE
#| fig-align: center

energy_weather |> 
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.4) + 
  geom_smooth() + 
  facet_wrap(vars(State))
```


## Catch

<br>

:::callout-warning
## Catch 

The previous example isn't a good use of data-density - lot's of empty space.

We aren't showing our data well either - trends by state aren't clear

:::

## Fixed or Free Scales 

<br>

:::callout-tip 

* We could let the scales differ, or be "free"

* But... humans compare things well on a common scale

* And... we can accidentally create misleading plots if we use different scales

It's a balancing act!

::: 

## Free Scales in `facet_wrap`

This is definitely better

```{r}
#| warning: FALSE
#| message: FALSE
#| fig-align: center

energy_weather |> 
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.4) + 
  geom_smooth() + 
  facet_wrap(vars(State), scales = "free") 
```

## Code 

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: TRUE
#| eval: FALSE

energy_weather |> 
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.4) + 
  geom_smooth() + 
  facet_wrap(vars(State), scales = "free") 
```

## Your turn 

<br>

::: callout-caution 

* Filter to the State of Victoria (`VIC`)

* Plot the `MaxTemp` against `Demand`

* But wrap by the `Day` of the week

::: 

## What I want 

```{r}
#| warning: FALSE
#| message: FALSE
#| fig-align: center

energy_weather |> 
  filter(State == "VIC") |>
  filter(!is.na(Day)) |>
  mutate(Day = factor(Day, 
                      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  facet_wrap(vars(Day), ncol = 2)
```

## Tricky!

:::callout-warning 

You may find your plot looks different to mine.

* The facet order is different 

* There is missing data

* And a different numbers of rows

This is where Generative AI is super useful to help you understand how to change each of these parts. 

::: 


## `facet_grid` 

::: callout-note
## `facet_grid`

Sometimes you want to display facets using two categorical variables 
  e.g. State and Day 
  
Notes: Use `~` to indicate a relationship between variables in R 
  e.g. `Day ~ State`
::: 

. . .

Example use: 

```{r}
#| eval: FALSE
#| echo: TRUE 

energy_weather |> 
  filter(!is.na(Day)) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth() +
  facet_grid(Day ~ State)
```

## Demand by Day and State

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: FALSE
#| fig-align: center

energy_weather |> 
  filter(!is.na(Day)) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth() +
  facet_grid(Day ~State)
```

## Bit Much 

<br>

:::callout-caution 

## But 

We expect a difference in demand when people are at home and at work 

Let's try another way to visualise this

:::

## Weekday vs Weekend 

Let's make a new variable for weekday or weekend. 

Here I'm using `if_else`.

```{r}
#| eval: FALSE
#| echo: TRUE 

energy_weather |> 
  filter(!is.na(Day)) |>
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) 
```

## Weekday or Weekend Demand by State

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: FALSE
#| fig-align: center

energy_weather |> 
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth() +
  facet_grid(Day_type ~ State, scales = "free")
```

## Your turn 

<br>

::: callout-caution 

* Run my code to create the new variable for weekday and weekend

* Plot the `MaxTemp` against `Demand`

* Create a grid that shows `Day_type` and `State`

::: 

## Best Version 

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: FALSE
#| fig-align: center

energy_weather |> 
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth(aes(group = Day_type, col = Day_type)) +
  facet_wrap(vars(State), scales = "free")
```

## Best Version 

:::callout-tip
## Why can't we have both? 

Using `group` and `facet_wrap` seems to show:  

* the weekday/weekend relationship best, and   

* show the differences between states

:::

 . . . 

:::callout-warning
## Subtle point 

`Scales = "free"` allows for different scales for each category. 

This **doesn't** mean each plot will have it's one x and y scale.

::: 

# Higher Dimensions {background-color="#006DAE"}

## Pairs plot

:::: callout-note
## Pairs plot

::: incremental
-   A pairs plot gives an array of plots.

-   This can be implemented using the *ggpairs* function in the *GGally* package.

-   You will get different plots depending on the type of variables in your data set

- Look at the [examaples here](https://r-charts.com/correlation/ggpairs/)

:::
::::

. . .

Example use: 

```{r, eval=FALSE, echo = TRUE}
library(GGally)
ggpairs(energy_weather)
```

# Combining Plots {background-color="#006DAE"}

## Patchwork 

::: callout-tip 
## Patchwork 

Sometimes we need plot panels that don't share a common variable or geometry

For this we can use [patchwork](https://patchwork.data-imaginist.com/articles/guides/layout.html)

Look at the patchwork documentation for more customisable layouts
  
Show you some basic ones

:::

## Some example plots

```{r}
#| echo: TRUE 

tas_data = data = energy_weather |>
         filter(State == "TAS")

temp_vs_demand_plot <- ggplot(tas_data) + 
  geom_point(aes(x = MaxTemp, y = Demand)) 

demand_distrib_plot <- ggplot(tas_data) +
  geom_density(aes(y = Demand)) 

weather_distrib_plot <- ggplot(tas_data) +
  geom_density(aes(x = MaxTemp))

```

## Example - side by side 

```{r}
#| echo: FALSE
#| fig-align: center 
library(patchwork)
temp_vs_demand_plot + demand_distrib_plot +
  plot_layout(widths = c(3, 1))
```

## Example - side by side 

```{r}
#| echo: TRUE
#| eval: false

library(patchwork)
temp_vs_demand_plot + demand_distrib_plot +
  plot_layout(widths = c(3, 1))
```

## Example - top and bottom

```{r}
#| fig-align: center
temp_vs_demand_plot / weather_distrib_plot + 
  plot_layout(heights = c(3, 1))
```

## Example - top and bottom

```{r}
#| echo: TRUE
#| eval: FALSE

temp_vs_demand_plot / weather_distrib_plot + 
  plot_layout(heights = c(3, 1))
```
# Wrap Up {background-color="#006DAE"}

## Summary

:::: callout-note
## What we've covered

**Today has been all about how create panels of plots**

::: incremental
-   We've learnt how use the `group` aesthetic 

-   We learnt how to create small multiples using facets
    -   `facet_wrap()`
    -   `facet_grid()`
    
<!-- -   We learnt how to create quick visual summaries -->
<!--     -   Displaying multiple dimensions -->
<!--     -   Using the pairs plot (`ggpairs`) -->
    
-   Learnt how to combine multiple plots together with `patchwork`
    
:::
::::

. . . 

:::callout-important
## Key Message 

There are lots of ways to visual data! Consider what comparisons are important.

:::

# Solutions {background-color="#006DAE"}

## Answers 

```{r}
#| echo: TRUE
#| eval: false 

ggplot(data = energy_weather, 
       aes(x = MaxTemp, y = Demand, col = State, group = State)) + 
  geom_point() + 
  geom_smooth() 
```

## Answers 

```{r}
#| echo: TRUE
#| eval: false 

energy_weather |> 
  filter(State == "VIC") |>
  filter(!is.na(Day)) |>
  mutate(Day = factor(Day, 
                      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  facet_wrap(vars(Day), ncol = 2)
```

## Answers 

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: TRUE
#| eval: FALSE

energy_weather |> 
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth() +
  facet_grid(Day_type ~State, scales = "free")
```

## Answers

```{r}
#| warning: FALSE
#| message: FALSE
#| echo: true
#| eval: false

energy_weather |> 
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) |>
  ggplot(aes(x = MaxTemp, y = Demand)) +
  geom_point(alpha = 0.25) + 
  geom_smooth(aes(group = Day_type, col = Day_type)) +
  facet_wrap(vars(State), scales = "free")
```


