---
title: "Tutorial-09"
author: Kate Saunders
format: 
  html:
    toc: true
    css: tutorial.css
    embed-resources: true
  pdf:
    toc: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'center')
show_solutions = FALSE
```

## Quarto Basics and Scrollytelling

### Learning Objectives

-   Practice communicating your results using Quarto

    -   Including practicing combining text, code, images and visualisations in the same document

-   Practice communicating your key messages by writing about your plots

-   Practice converting your Quarto into a storytelling document

### Preparation

-   We will use the energy and weather data again, and the plots we created in Workshop 6.

```{r echo = TRUE, eval = FALSE}
library(tidyverse)
library(here)
energy = read_csv("data/energydata.csv")
weather = read_csv("data/weather.csv")
energy_weather = full_join(energy, weather, by = c("Date", "State"))

energy_weather |> 
  mutate(
    Day_type = if_else(Day %in% c("Sat", "Sun"),
                       "Weekend", "Weekday")
  ) |>
  ggplot(aes(x = MaxTemp, y = Demand, group = Day_type, col = Day_type)) +
  geom_point(alpha = 0.25) + 
  geom_smooth() +
  facet_wrap(vars(State), scales = "free")
```

### Task 1

i.  Create a Quarto document that contains the above plot (don't show your code)

ii. Write a sentence above the plot about the key message

iii. Below the plot write about the features of the plots and the key aesthetic mappings

iv. Write about any of the secondary messages or other important details

v.  Write about any important caveats

:::callout-important

A `.qmd` file you can download that provides a solution for this example is on the Moodle - Day 9 Wrap Up.

:::

### Task 2

Turn your Quarto into a scroll based story.

i.  Create a copy of your Task 1 Quarto file.

ii. Change the format to a `closeread-html`

iii. Make the plot a sticky object by giving it an ID (for example `:::{#cr-myplot}`)

iv. Add the headers `:::{.cr-section}` for where you want the scrolling to start and `:::` for where you want the scrolling to finish

v.  Inside your scrolly section, add the trigger for when you want your plot to appear and make the messages scroll by one at a time

vi. Advanced: Add in other plots to improve the telling of this data story

:::callout-important

A `.qmd` file you can download that provides a solution for this example is on the Moodle - Day 9 Wrap Up.
 

Remember this file won't run unless you've correctly installed the `closeread` template in the same directory as your `.qmd` file.

:::
