---
title: "Tutorial-10"
author: Dr. Kate Saunders and Krisanat Anukarnsakulchularp
format: 
  html:
    toc: true
    css: tutorial.css
    embed-resources: true
  pdf:
    toc: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'center')
show_solutions = FALSE
```

## Exploratory Visualisation

### Learning Objectives

-   Practice communicating the key messages of your visualisation 

-   This will require iterating on your visualisation 

- You will add narrative elements such as text annotations to make your key messages clear

### Preparation

-   Come ready with your favourite plot(s) from Assignment 1.

- Download the `crime_statistics.csv` from Moodle 

- Run the following code to produce the below plot

```{r message=FALSE}
library(tidyverse)
library(colorblindr)
library(here) 

crime_data <- read_csv(here("data/crimestatistics.csv"))

monash_crime_data <- crime_data |> 
  filter(local_government_area == "Monash",
         year == 2024) |> 
  group_by(offence_division) |> 
  summarise(total = sum(incidents_recorded)) 


monash_crime_plot <- ggplot(
  data = monash_crime_data, 
  aes(x = offence_division, y = total, fill = offence_division)) +
  geom_col() +
  labs(y = "Total number of incidences", x = "", 
       title = "Total number offences in the Monash LGA in 2024") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0.1),
        legend.title = element_blank()) +
  colorblindr::scale_fill_OkabeIto() 

monash_crime_plot
```
### Motivation

Effective communication often requries extending you visualisation by adding narrative elements to guide your audience. This includes elements such as text, highlighting, arrows and labels.

In this tutorial, we'll practice adding narrative elements to our visualisation, so we can better deliver the key messages to the audience.

### Exercise 1

Pick your best visualisation from Assignment 1 and write down the key messages. Show this visualisation to the person next to you. Ask them to identify the key messages from the visualisation alone. 

Are the key messages they identified the same as what you wrote down?

:::callout-warning 
## Think about your audience 

It can be tempting to assume that it is the audience's fault for not understanding your key messages. They weren't smart enough, they just didn't get it etc. 

The reality this just doesn't matter. Also it is almost never the case. You simply need to do better! 

If your test audience didn't understand your key messages you need to critically reflect. How you could make the messages clearer or simpler?

:::

### Exercise 2

Add a text box to the plot from your assignment to highlight that key message.

Here is some example code you might find useful.

```{r}
monash_crime_plot + 
  annotate("label", x = 4.5, y = 6000, label = "Category B has the \n highest number of incidenes")
```

::: callout-caution
## Question

Is adding a text box alone enough to convey the messages? Discuss this with the person next to you.
:::

### Exercise 3

Now, let's add an arrow to highlight elements the plot from your assignment.

Again here is some example code to get you started:

```{r}

monash_crime_plot +
  annotate("segment", xend = 6, yend = 2500, x = 5, y = 42500, arrow = arrow(length = unit("3", "mm"))) + 
  annotate("label", x = 4.5, y = 40000, label = "Category F has the \n lowest number of incidenes") +
  ylim(c(0, 45000))
```

*You may like to note the order of the layers and I have changed the range of the y axes.*

::: callout-caution
## Task

Show your plot to the person next to you and see if they have received your key message from the infographic you created.
:::

### Finishing Up 

* Make sure you understand how to position text or a label on your plot by changing the x and y coordinates. 

* Makes sure you understand how to add a line segment with an arrow.

* Think about how you might use colour to further highlight these messages.

* Lastly, don't always go adding narrative elements. Think about if narrative elements are appropriate given the medium and if they enhance your key messages.

:::callout-important 
## Assignment 2 

Notice the Monash Crime plot is polished:

* The title, axes labels and legends have all been editted to be more interpretable. 

* The text is appropriately sized.

* The colour scale is colour blind friendly.  

**We expect this level of polish from you on Assignment 2.**

:::


##### Material developed by Dr. Kate Saunders and Krisanat.Anukarnsakulchularp.

##### © Copyright 2024 Monash University
