---
title: "Tutorial-06"
author: Kate Saunders
format: 
  html:
    toc: true
    css: tutorial.css
    embed-resources: true
  pdf:
    toc: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'center')
show_solutions = FALSE
```

## Visualisation in R: Facets and Group Aesthetic Mapping

### Learning Objectives

-   Practice using small multiples to visualise your data

- This will involve using `facet_wrap` and `facet_grid`

- Also practice grouping your data by categorical variables for visualisation

### Preparation

-   We expect you to be using an R project for all tutorials

-   Download the dataset from Moodle, `boston_celtics.csv,` and place it in a folder called `data` within your R Project.

### Tasks

Building on tutorial 4, you will be creating visualisations in `ggplot2` to analyse sporting statistics from the Boston Celtics NBA basketball team.

```{r echo = TRUE}
library(tidyverse)
library(here)
boston_celtics <- read_csv(here("data", "boston_celtics.csv"))
```

### Task 1

Using `facet_wrap` recreate the following plot:

```{r echo = show_solutions}
ggplot(data = boston_celtics) + 
  geom_histogram(aes(x = team_score), 
                 fill = "forestgreen", alpha = 0.4, binwidth = 4) +
  geom_vline(aes(xintercept = mean(team_score)), linetype = "dotted") +
  facet_wrap(~season, ncol = 2) + 
  theme_bw() + 
  xlab("Team Score") + 
  ylab("Count") +
  ggtitle("Boston Celtics Past Season Performance")
```

**Hint:** Look up `geom_vline()` to add the mean score each season to the plot.

### Task 2

Using `facet_grid` recreate the following plot:

```{r}
#| echo: false
#| eval: true

boston_celtics = boston_celtics |>
  mutate(team_winner = as.factor(team_winner)) |>
  mutate(
    team_winner = 
      recode(team_winner,
             "TRUE" = "Won", "FALSE" = "Lost"),
    team_home_away = 
      recode(team_home_away,"home" = "Home", "away" = "Away"))
```

```{r}
#| echo: false
#| eval: true
ggplot(data = boston_celtics) + 
  geom_boxplot(aes(x = team_score),
                 fill = "forestgreen", alpha = 0.4, binwidth = 1) +
  facet_grid(team_home_away ~ team_winner) + 
  theme_bw()
```

**Hint:** What's tricky here is changing the order (levels) of the categorical variables. The default for how categorical variables are displayed on panels of plots is always alphabetical.

```{r}
#| echo: TRUE
#| eval: FALSE
boston_celtics = boston_celtics |>
  mutate(
    team_winner = 
      factor(team_winner, levels = c("Won", "Lost")),
    team_home_away = 
      factor(team_home_away, levels = c("Home", "Away"))
  )
```
We need to: 

* Tell R it's a categorical variable we make it a `factor` type

* Change the order of the levels for plotting using `recode`.

[Here](https://rstudio.github.io/cheatsheets/html/factors.html) is a cheat sheet on `factors` you can download to help you. 

### Task 3

Use the `group` aesthetic mapping and `facet_wrap` to create the following plot: 

```{r}
ggplot(boston_celtics) + 
  geom_density(aes(x = team_score, group = team_home_away, col = team_home_away), linewidth = 1.1) + 
  scale_colour_manual(label = c("Home", "Away"), values = c("Home" = "forestgreen", "Away" = "lightgreen")) +
  facet_wrap(vars(team_winner)) +
  theme_bw() + 
  xlab("Team Score") + 
  ylab("Density") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

```
**Hint:** You can adapt the colour scale from Tutorial 4 and the `linewidth` is increased from default.

##### © Copyright 2025 Monash University
